A mechanical Model of Rod-Shaped Bacteria

### 2025-06-23 paper/main.tex
add plot for overview of project

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{float}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{onimage}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage{tikz}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{acronym}

\usetikzlibrary{trees}

@@

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{float}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{onimage}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage{tikz}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{acronym}
\usepackage{xcolor}

\usetikzlibrary{trees}
\usetikzlibrary{shapes.geometric, arrows}

\definecolor{color1}{HTML}{6bd2db}
\definecolor{color2}{HTML}{0ea7b5}
\definecolor{color3}{HTML}{0c457d}
\definecolor{color4}{HTML}{ffbe4f}
\definecolor{color5}{HTML}{e8702a}

@@

\begin{figure}
    \centering
    \tikzstyle{rect} = [
        rectangle,
        % rounded corners,
        minimum width=4cm,
        minimum height=1cm,
        text centered,
        draw=black,
        fill=color4!70,
    ]
    \tikzstyle{frect} = [rect, fill=color3!40]
    \tikzstyle{arrow} = [thick,->,>=stealth]
    \tikzstyle{darrow} = [arrow, dashed]
    \begin{tikzpicture}
        \node (model) [rect, align=center] at (0,0) {Mathematical \&\\ Computational Model};
        \node (extraction) [rect] at (6, 0) {Data Extraction};
        \node (params) [rect] at (6, -2) {Parameter Estimation};
        \node (visualization) [rect] at (0, -2) {Visualization};

        \draw[arrow] (model.east) -- (extraction.west);
        \draw[darrow] (params.north) -- node[anchor=south, rotate=-10] {validate} (model.south);
        \draw[arrow] (model.south) -- (visualization.north);
        \draw[arrow] (visualization.east) -- (params.west);
        \draw[arrow] (extraction.south) -- (params.north);

        % Future stuff
        \node (realimages) [frect] at (0, -4) {Data Generation};
        \node (segmentation) [frect] at (6, -3.5) {Cell Segmentation};
        \node (tracking) [frect] at (6, -4.5) {Cell Tracking};

        \draw[arrow] (visualization.south) -- (realimages.north);
        \draw[arrow] (realimages) -- (segmentation.west);
        \draw[arrow] (realimages) -- (tracking.west);
    \end{tikzpicture}
    \caption{
        The project can be divided into 4 subparts (yellow) that interact with and depend on each
        other.
        Additional blocks (blue-gray) indicate possible future extensions of this project.
        Arrows with solid lines indicate dependencies between the subparts while dashed arrows carry
        special meaning, which is specifically annotated.
        The mathematical models provides the theoretical framework and is thus the parent of all
        following work.
        "Data Extraction" allows us to read variable and parameter values directly from microscopic
        images.
        The obtained data can then be used to initialize a numerical simulations and thus estimate
        parameters.
        On the other hand, results produced by our model can be visualized as 3D renders and
        projected to cell-masks.
        These algorithms can be exploited to construct cost-functions which can be used in the
        parameter estimation process.
        In future work, this visualization scheme could be extended to generate realistic
        microscopic images with cell-masks and thus train cell-segmentation and cell-tracking
        algorithms.
    }
    \label{fig:flowchart-project-structure}
\end{figure}
